Title

### 2016-04-26 paper/sections/Overview.tex
More text on S3

\lstinputlisting[linerange=11-26]{../src/paper/Circuit.java}%APPLY:INIT
We change the implementation language constructs from classes to interfaces and
their fields to unimplemented getter methods. The purpose of this modification
is to allow types of the fields to be refined in future, which is vital for
retaining extensibility.

\subsection{Adding New Syntax}
Shallow embedding makes it easy to add new syntax, both in FP and OOP.
Suppose that we would like to add two new constructs to the language:
the first is \emph{stretch ns c}, which inserts additional wires into the circuit \emph{c} by summing up \emph{ns};
the second is $above c_1 c_2$, which combines two circuits of the same width vertically.
With these new constructs, more complex circuit can be constructed.
For example, Figure~\ref{} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.

To accomplish this goal in Haskell, we can simply define two more cases,
\texttt{stretch} and \texttt{above}, for the semantic function:
\lstinputlisting[linerange=59-60]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
Similarly, defining two new interfaces
\texttt{Stretch} and \texttt{Above} that both extend \texttt{Circuit} and
implement the \texttt{width} method is all we need to do in Java:

\lstinputlisting[linerange=29-40]{../src/paper/Circuit.java}%APPLY:SYNTAX

@@

Adding new semantics, however, becomes hard for shallow embedding.
Gibbons and Wu worked around this problem in the following way:
\lstinputlisting[linerange=47-51]{../src/paper/Circuit.java}%APPLY:HS_SEMANTICS

which is similar to how we add \texttt{desugar} to the implementation in Section~\ref{}.
The difference is that they use a tuple instead of a record to merge the
semantic domains, then define the two semantic functions simultaneously for each
case, and split the definitions through projections on the tuple in the end.
This solution, however, modifies existing code, breaking the requirement of EP.

Conversely, the support of covariant type-refinements and inheritance for OO
languages allows us to add new semantics in a \emph{modular} way:
%Different from records or tuples, interfaces are extensible.
\lstinputlisting[linerange=47-75]{../src/paper/Circuit.java}%APPLY:SEMANTICS

@@

\subsection{\name's Support}
There exists some boilerplate for the Java solution presented above:
\begin{itemize}
  \item Interfaces that represent language constructs should be instantiated for
    creating objects;
  \item Type-refinements should be done manually
  \item The inheritance relation needs to be repeatedly stated for all interfaces.
\end{itemize}
Also, programmers would not get warned if they forget to extend all the
constructs with new semantics.

\name addresses all these problems through embracing family polymorphism and code instrument.
By using \name, we can refactor the extensions in the way shown in Figure\ref{}.
\lstinputlisting[linerange=80-83]{../src/paper/Circuit.java}%APPLY:FAMILY
\lstinputlisting[linerange=112-122]{../src/paper/Circuit.java}%APPLY:FAMILY_SYNTAX
\lstinputlisting[linerange=126-141]{../src/paper/Circuit.java}%APPLY:FAMILY_SEMANTICS

Note that \name generates \texttt{of} method for class-like interfaces for
constructing objects. The static method \texttt{of} returns an instance of an anonymous class
that implements the interface with all
getters implemented and its instance is returned.
Moreover, \name has a good support family polymorphism.
Client users need only explicitly declared the dependencies on other families
through (...). \name will re-declare all members from the base families in the extended family.
And the dependencies between members of families are automatically deducted and
implicitly expressed in member re-declaration.

>>>>

\lstinputlisting[linerange=11-26]{../src/paper/sec3/Circuit.java}%APPLY:INIT
We change the implementation language constructs from classes to interfaces and
their fields to unimplemented getter methods. The purpose of this modification
is to allow types of the fields to be refined in future, which is vital for
retaining extensibility.

\subsection{Adding New Syntax}
Shallow embedding makes it easy to add new syntax, both in FP and OOP.
Suppose that we would like to add two new constructs to the language:

\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
summing up \emph{ns} and $above c_1 c_2$ combines two circuits of the same width vertically.
With these new constructs, more complex circuits can be constructed.
For example, Figure~\ref{} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.

To accomplish this task in Haskell, we can simply define two more cases,
\texttt{stretch} and \texttt{above}, for the semantic function:
\lstinputlisting[language=haskell,linerange=45-46]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
Similarly, defining two new interfaces
\texttt{Stretch} and \texttt{Above} that both extend \texttt{Circuit} and
implement the \texttt{width} method is all we need to do in Java:

\lstinputlisting[linerange=29-40]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX

@@

Adding new semantics, however, becomes hard for shallow embedding.
Gibbons and Wu worked around this problem in the following way:
\lstinputlisting[language=haskell,linerange=2-6]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
which is similar to how we add \texttt{desugar} to the implementation in Section~\ref{}.
The difference is that they use a tuple instead of a record to merge semantic
domains, then define the two semantic functions simultaneously for each
case, and split the semantic functions through projections on the tuple in the end.
This solution, however, modifies existing code, breaking the requirement of EP.

Conversely, the support of covariant type-refinements and inheritance for OO
languages allows us to add new semantics in a \emph{modular} way:
%Different from records or tuples, interfaces are extensible.
\lstinputlisting[linerange=47-75]{../src/paper/sec3/Circuit.java}%APPLY:SEMANTICS

@@

\subsection{\name's Support}
There exists some boilerplate for the Java solution presented above:
\begin{itemize}
  \item Interfaces that represent language constructs should be instantiated for
    creating objects;
  \item Type-refinements should be done manually;
  \item Similar inheritance relationships have to be repeatedly stated for all interfaces.
\end{itemize}
Worse still, programmers would not get warned if they forget to extend all the
constructs with new semantics.

\name addresses all these problems through embracing family polymorphism and code instrument.
By using \name, we can refactor the extensions in the way shown in Figure\ref{}.
\lstinputlisting[linerange=80-82]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY
\lstinputlisting[linerange=110-112]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SYNTAX
\lstinputlisting[linerange=131-156]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SEMANTICS

Note that initially all the definitions are put inside a \texttt{Family} interface.
Extensions are defined inside another interface that extends base families.
For syntax extension, we just move the definitions of new constructs into
\texttt{SyntaxExt} interface.
Semantic extension is more interesting, let us look at it in detail.
Since all the definitions are nested interfaces, their names are local and can
be reused in other families. The inheritance relationships is stated only once at family level, then \name can infer the inheritance relationships for members inside the family according to their names. Also, there is no need to
manually refine the return type, as it is done by \name. Moreover, a static \texttt{of}
method would be generated for class-like interfaces, serving as the constructor
for the interface. For example, the \texttt{Beside} inside \texttt{Semantics} after code instrumentation will look like this:

\lstinputlisting[linerange=164-176]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
% TODO: not implemented yet
Family polymorphism on only gives more safety on the client code but also helps
language implementers catch bug early.
For example, if one forgets to implement, say cases from \texttt{Family}, in
\texttt{Semantics}, she would get warned they have not implemented the
\texttt{wellSized} method because class-like interfaces should not contain any
unimplemented methods except for getters.
These errors are captured by automatically re-declaring members with inheritance relationships.
%The \texttt{of} method is  returns an instance of an anonymous class
%that implements the interface with all getters implemented and its instance is returned.

### 2016-04-26 paper/sections/ShallowOO.tex
More text on S3

\lstinputlisting[language=haskell,linerange=2-5]{./code/shallowCircuit.hs}%APPLY:CTYPE
where \texttt{Circuit} models the semantic domain and should be later filled in with a
concrete type according to the semantics of the language.
For example, if the semantics of \dsl is to calculate the width of a
circuit, the definition would be:
\lstinputlisting[language=haskell,linerange=11-16]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1

@@

We continue rewriting the definition. This time, we use a record to represent
the result and give it a name:
\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
The implementation is apparently shallow since it is no different from the last version.
At this stage, however, the definition looks so like an OO program that we can easily
translate it into an OO language like Java:
\lstinputlisting[linerange=5-24]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
The record type maps to an interface with its field declared as an abstract method.
Each case corresponds to an class and its parameters become fields of the class.
And the class implements the \texttt{Circuit} interface by giving \texttt{width}
a definition. This implementation, in fact, is an instance of the \interp pattern.
In other words, shallow embedding and the \interp are essentially the same thing.
It may also worth mentioning that deep embedding is closely related to the
\textsc{Visitor} pattern.

@@

We can implement this desugaring operation in Java:
\lstinputlisting[language=haskell,linerange=29-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
The return type of \texttt{desugar} is \texttt{Circuit} reflecting that it is a
transformation. The implementation is straightforward: \emph{fan n} just returns
itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wrap back
their results into an instance of \texttt{Beside}; \emph{identity n} encodes the above
formula.

@@

\lstinputlisting[language=haskell,linerange=40-55]{./code/shallowCircuit.hs}%APPLY:DESUGAR
The definition is a little bit tricky since it makes use of Haskell's laziness.

>>>>

\lstinputlisting[language=haskell,linerange=2-2]{./code/shallowCircuit.hs}%APPLY:CIRCUIT_TYPE
\lstinputlisting[language=haskell,linerange=5-7]{./code/shallowCircuit.hs}%APPLY:TYPES
where \texttt{Circuit} models the semantic domain and should be later filled in with a
concrete type according to the semantics of the language.
For example, if the semantics of \dsl is to calculate the width of a
circuit, the definition would be:
\lstinputlisting[language=haskell,linerange=11-16]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1

@@

We continue rewriting the definition. This time, we use a record to represent
the result and give it a name:
\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
The implementation is apparently shallow since it is no different from the last version.
At this stage, however, the definition looks so like an OO program that we can easily
translate it into an OO language like Java:
\lstinputlisting[linerange=5-24]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
The record type maps to an interface with its field declared as an abstract method.
Each case corresponds to an class and its parameters become fields of the class.
And the class implements the \texttt{Circuit} interface by giving \texttt{width}
a definition. This whole implementation is, in fact, an instance of the \interp
pattern and the combinator class \texttt{Beside} employs the \textsc{Composite} pattern.
In other words, shallow embedding and the \interp are essentially the same thing.
It may also worth mentioning that deep embedding is closely related to the
\textsc{Visitor} pattern.

@@

We can implement this desugaring operation in Java:
\lstinputlisting[linerange=29-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
The return type of \texttt{desugar} is \texttt{Circuit} reflecting that it is a
transformation. The implementation is straightforward: \emph{fan n} just returns
itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wrap back
their results into an instance of \texttt{Beside}; \emph{identity n} encodes the above
formula.

@@

\lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
The definition is a little bit tricky since it makes use of Haskell's laziness.
